**Socio-economic and geographic determinants of Covid-19 vaccination**

### 2023-04-23 manuscript_v4/02-methods.qmd
Started v4 to implement final changes before submission

We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences (henceforth Fields) and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto, and which ran between September 30th, 2021 and January 17th,2022. The survey collected socio-economic information from participants (@tbl-covariates), recorded their location (using the nearest municipality), and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no". The original dataset contained 39,029 entries (where each entry corresponded to a unique respondent).

This dataset was cleaned to remove outliers that were identified during preliminary analyses, and processing the geographical information in the survey (city where the survey was responded) in order to match each city to its correspondent Health Region.

@@

Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions. The low representation of these Health Regions in the dataset is noticeable in @fig-map,which shows that responses from these areas came from a relatively low number of cities when compared to the most populated Health Regions, such as the Toronto or Central Regions. We omitted the North East and North East Health Regions from further analyses due to the low number of entries. Therefore, the total number of unique entries used for analysis was 3,551 which included the East, Central, Toronto, and West Health Regions and that covered the period between September 30th,2021 and December 12th, 2021.

@@

We used a logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @tbl-covariates, the Health Regions from Ontario indicated in @sec-data, and the interactions between Race and Health Region, and Race and income, as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].

The model was fitted first to the clean dataset to obtain uncorrected estimates. Additionally, because we identified differences between the proportions of all the socio-economic factors included in the analysis (@tbl-covariates) and the Census data for Ontario, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using Census socio-economic data and Health Region population totals, in order to obtain corrected estimates from the model. Details regarding the correction can be found in the Appendix. 
All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], and `quarto` [@quarto].

>>>>

We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences (henceforth Fields) and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto, and which ran between September 30th, 2021 and January 17th,2022. The survey collected socio-economic information from participants (@tbl-covariates), recorded their location (using the nearest municipality), the date of access to the survey, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no". The original dataset contained 39,029 entries (where each entry corresponded to a unique respondent).

This dataset was cleaned to remove outliers that were identified during preliminary analyses, and the geographical information in the survey (city where the survey was responded) was used to match each city to its correspondent Health Region.

@@

Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions. The low representation of these Health Regions in the dataset is noticeable in @fig-map, which shows that responses from these areas came from a relatively low number of cities when compared to the most populated Health Regions, such as the Toronto or Central Regions. We omitted the North East and North East Health Regions from further analyses due to the low number of entries. Therefore, the total number of unique entries used for analysis was 3,549 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and December 12th, 2021.

@@

We used a logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @tbl-covariates, the month when the survey was answered, the Health Regions from Ontario indicated in @sec-data, and the interactions between Race and Health Region, and Race and income, as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].

The model was fitted first to the clean dataset to obtain uncorrected estimates. Because we identified differences between the proportions of all the socio-economic factors included in the analysis (@tbl-covariates) and the Census data for Ontario, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using Census socio-economic data and Health Region population totals, in order to obtain corrected estimates from the model. Details regarding the correction can be found in the Appendix. 
All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], and `quarto` [@quarto].

### 2023-04-23 manuscript_v4/03-results.qmd
Started v4 to implement final changes before submission

Within the age groups of survey respondents, the age group between 16-34 years had the highest representation in the survey responses (1,520 or 42.8\% of all responses). Within this age bracket, 73\% of respondents indicated having received the vaccine, whereas the lowest vaccination rate was in the bracket of those 55 years of age and above, with 72\%. The Health Region with highest representation in the survey was Toronto, accounting for 1,323 entries (37.3\% of the total), with a vaccination rate of 72\%. Regarding race/ethnicity, individuals that identified as White/Caucasian represented 37\% of all entries (1,312) and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine. On the other hand, the ethnic group with the lowest number of entries in the survey was Latin American, with a total of 180, or 5\% of all entries. Vaccination rates across all minority groups were below the value reported by White/Caucasians, with the lowest vaccination rate (60\%) being reported by individuals that identified as Indigenous.

```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination Status)"
#| message: false
#| echo: false

@@

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Month,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income")
  ) %>% bold_labels()

@@

t0<-m0 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                            Age_group_ord ~ "Age Group",
                            income_ord="Income")) %>%
  bold_labels()

t1<-m1 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                            Age_group_ord ~ "Age Group",
                            income_ord="Income")) %>%
  bold_labels()

t2<-m2 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                            Age_group_ord ~ "Age Group",
                            income_ord="Income")) %>%
  bold_labels()

>>>>

Within the age groups of survey respondents, the age group between 16-34 years had the highest representation in the survey responses (1,520 or 42.8\% of all responses). Within this age bracket, 73\% of respondents indicated having received the vaccine, whereas the lowest vaccination rate was in the bracket of those 55 years of age and above, with 72\%. The Health Region with highest representation in the survey was Toronto, accounting for 1,323 entries (37.3\% of the total), with a vaccination rate of 72\%. Across the three months covered by the survey, the highest number of entries was in October of 2021, with a total of 1,732 respondents, which corresponded to 49\% of all the survey entries. Additionally, the proportion of respondents who reported being vaccinated was similar in all months, ranging between 72\% and 75\%, with the highest value corresponding to December of 2021.

Regarding race/ethnicity, individuals that identified as White/Caucasian represented 37\% of all entries (1,312) and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine. On the other hand, the ethnic group with the lowest number of entries in the survey was Latin American, with a total of 180, or 5\% of all entries. Vaccination rates across all minority groups were below the value reported by White/Caucasians, with the lowest vaccination rate (60\%) being reported by individuals that identified as Indigenous.

```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination Status)"
#| message: false
#| warning: false
#| echo: false

@@

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Month,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income (CAD)")
  ) %>% bold_labels()

@@

t0<-m0 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                            Age_group_ord ~ "Age Group",
                            income_ord="Income (CAD)")) %>%
  bold_labels()

# t1<-m1 %>%
#   tbl_regression(exponentiate = TRUE,
#                  label=list(Health_Region ~ "Health Region",
#                             Age_group_ord ~ "Age Group",
#                             income_ord="Income")) %>%
#   bold_labels()

t2<-m2 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                            Age_group_ord ~ "Age Group",
                            income_ord="Income (CAD)")) %>%
  bold_labels()
